---
title: "shiny_fundamentals_with_r"
author: "Salerno"
date: "02/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Shiny Fundamentals with R

Learn how to use Shiny, a popular R package, to build highly interactive web applications—without needing any other programming languages. Using Shiny you can share your analyses as dashboards and visualizations to bring your data to life.

In this track, you’ll get hands-on with Gapminder, Star Wars, and NASA datasets, as you discover how to represent raw data using explorable visualizations including plots and word clouds. You’ll also improve your ability to make discoveries and communicate these visually to non-technical stakeholders—an invaluable skill for any data analyst or data scientist!


## 1) Building Web Applecations with Shiny in R

Shiny is an R package that makes it easy to build highly interactive web apps directly in R. Using Shiny, data scientists can create interactive web apps that allow your team to dive in and explore your data as dashboards or visualizations. If you want to bring your data to life, Shiny is the way to go! Using data about baby names, food ingredients, and UFO sightings, you'll build a variety of different Shiny apps that leverage different inputs and outputs. You’ll also learn the basics of reactive expressions. By the end of this course, you’ll have the Shiny skills you need to build your first app in R.

### 1.1) Get started with Shiny

To kick off the course you'll learn what a web app is and when you should build one, plus build a few apps of your own! You'll first learn to make text inputs and outputs in a few ways, including exploring the popularity of certain names over time.

### 1.2) "Hello, World" app input (UI)

The "Hello, World!!!" app, while enthusiastic, could be more fun. Extend the app and have it wish a hello to a specific person. A user will enter a name and the app will update to wish that name "Hello".

For users to enter text, you'll have to use a text-specific shiny input function to capture it. Recall that shiny makes available a number of input functions depending on what kind of input you'd like to capture from your users.

```{r}

library(shiny)

ui <- fluidPage(
	textInput("name", "Enter the name:")
  
)

server <- function(input, output) {
  
}

shinyApp(ui = ui, server = server)

```

### 1.3) "Hello, World" app output (UI/Server)

To finish up your "Hello, world" app, you'll have to actually display the text that's input.

Recall this is how you construct an output from an input:

```{r}

ui <- fluidPage(
	textInput("name", "What is your name?"),
	# CODE BELOW: Display the text output, greeting
    # Make sure to add a comma after textInput()
  textOutput("greeting")
)

server <- function(input, output) {
	# CODE BELOW: Render a text output, greeting
  output$greeting <- renderText({
    paste("Hello, ", input$name)
  })
  
  
}

shinyApp(ui = ui, server = server)

```

### 1.4) Add input (UI)

This app will allow users to enter a baby name and visualize the popularity of that name over time.

The first step is to add a text input to the UI that will allow a user to enter their (or any other) name. Try using the optional default argument this time around.


```{r}

ui <- fluidPage(
  # CODE BELOW: Add a text input "name"
  textInput("name", "Enter your name", "David")
  
)
server <- function(input, output, session) {


}
shinyApp(ui = ui, server = server)

```

### 1.5) Add output (UI/Server)

The next step in building your app is to add an empty plot as a placeholder. Recall that in order to add a plot p assigned to an object named x to a Shiny app, you need to:

- Render the plot object using renderPlot({p}).
- Assign the rendered plot to output$x.
- Display the plot in the UI using plotOutput("x").

```{r}

library(ggplot2)

ui <- fluidPage(
  textInput('name', 'Enter Name', 'David'),
  # CODE BELOW: Display the plot output named 'trend'
  plotOutput("trend")
  
)
server <- function(input, output, session) {
  # CODE BELOW: Render an empty plot and assign to output named 'trend'
  output$trend <- renderPlot({
   trend <- ggplot() 
  })
}
shinyApp(ui = ui, server = server)

```

### 1.6) Update layout (UI)

You can use layout functions provided by Shiny to arrange the UI elements. In this case, we want to use a sidebarLayout(), where the input is placed inside a sidebarPanel() and the output is placed inside the mainPanel(). You can use this template to update the layout of your app.

sidebarLayout(
  sidebarPanel(p("This goes into the sidebar on the left")),
  mainPanel(p("This goes into the panel on the right"))
)

We have pre-loaded the shiny and ggplot2 packages for you. Note that p('hello') returns an HTML paragraph with the text "hello".

```{r}

ui <- fluidPage(
  titlePanel("Baby Name Explorer"),
  # CODE BELOW: Add a sidebarLayout, sidebarPanel, and mainPanel
  sidebarLayout(
    sidebarPanel(
      textInput('name', 'Enter Name', 'David')
      ),
    mainPanel(
    plotOutput('trend')
    )
  )
)

server <- function(input, output, session) {
  output$trend <- renderPlot({
    ggplot()
  })
}
shinyApp(ui = ui, server = server)

```

### 1.7) Update output (server)

You are almost there! The final step is to update the plot output to display a line plot of prop vs. year, colored by sex, for the name that was input by the user. You can use this plot template to create your plot:

ggplot(subset(babynames, name == "David")) +
  geom_line(aes(x = year, y = prop, color = sex))
  
Recall that a user input named foo can be accessed as input$foo in the server. We have already pre-loaded the shiny and ggplot2 packages, as well as the babynames dataset.

```{r}



ui <- fluidPage(
  titlePanel("Baby Name Explorer"),
  sidebarLayout(
    sidebarPanel(textInput('name', 'Enter Name', 'David')),
    mainPanel(plotOutput('trend'))
  )
)
server <- function(input, output, session) {
  output$trend <- renderPlot({
    # CODE BELOW: Update to display a line plot of the input name
    ggplot(subset(babynames::babynames, name == "David")) +
    geom_line(aes(x = year, y = prop, color = sex))
    
  })
}
shinyApp(ui = ui, server = server)

```

### 2.1) Add a select input

```{r}

library(shiny)
library(tidyverse)

ui <- fluidPage(
  titlePanel("What's in a Name?"),
  # CODE BELOW: Add select input named "sex" to choose between "M" and "F"
  selectInput('sex', 'Select Sex', choices = c("F", "M")),
  # Add plot output to display top 10 most popular names
  plotOutput('plot_top_10_names')
)

server <- function(input, output, session){
  # Render plot of top 10 most popular names
  output$plot_top_10_names <- renderPlot({
    # Get top 10 names by sex and year
    top_10_names <- babynames::babynames %>% 
      # MODIFY CODE BELOW: Filter for the selected sex
      filter(sex == input$sex) %>% 
      filter(year == 1900) %>% 
      top_n(10, prop)
    # Plot top 10 names by sex and year
    ggplot(top_10_names, aes(x = name, y = prop)) +
      geom_col(fill = "#263e63")
  })
}

shinyApp(ui = ui, server = server)

```

### 2.2) Add a slider input to select year

```{r}

ui <- fluidPage(
  titlePanel("What's in a Name?"),
  # Add select input named "sex" to choose between "M" and "F"
  selectInput('sex', 'Select Sex', choices = c("F", "M")),
  # CODE BELOW: Add slider input named 'year' to select years  (1900 - 2010)
  sliderInput('year', 'Select year', value = 1900, min = 1900, max = 2010),
  # Add plot output to display top 10 most popular names
  plotOutput('plot_top_10_names')
)

server <- function(input, output, session){
  # Render plot of top 10 most popular names
  output$plot_top_10_names <- renderPlot({
    # Get top 10 names by sex and year
    top_10_names <- babynames::babynames %>% 
      filter(sex == input$sex) %>% 
    # MODIFY CODE BELOW: Filter for the selected year
      filter(year == input$year) %>% 
      top_n(10, prop)
    # Plot top 10 names by sex and year
      ggplot(top_10_names, aes(x = name, y = prop)) +
        geom_col(fill = "#263e63")
  })
}

shinyApp(ui = ui, server = server)

```

### 2.3) Add a table output

```{r}

ui <- fluidPage(
  titlePanel("What's in a Name?"),
  # Add select input named "sex" to choose between "M" and "F"
  selectInput('sex', 'Select Sex', choices = c("F", "M")),
  # Add slider input named "year" to select year between 1900 and 2010
  sliderInput('year', 'Select Year', min = 1900, max = 2010, value = 1900),
  # CODE BELOW: Add table output named "table_top_10_names"
  tableOutput("table_top_10_names")
  
)
server <- function(input, output, session){
  # Function to create a data frame of top 10 names by sex and year 
  top_10_names <- function(){
    top_10_names <- babynames::babynames %>% 
      filter(sex == input$sex) %>% 
      filter(year == input$year) %>% 
      top_n(10, prop)
  }
  # CODE BELOW: Render a table output named "table_top_10_names"
  output$table_top_10_names <- renderTable({top_10_names()})
  
  
}
shinyApp(ui = ui, server = server)

```

### 2.4) Add an interactive table output

```{r}

ui <- fluidPage(
  titlePanel("What's in a Name?"),
  # Add select input named "sex" to choose between "M" and "F"
  selectInput('sex', 'Select Sex', choices = c("M", "F")),
  # Add slider input named "year" to select year between 1900 and 2010
  sliderInput('year', 'Select Year', min = 1900, max = 2010, value = 1900),
  # MODIFY CODE BELOW: Add a DT output named "table_top_10_names"
  DT::DTOutput('table_top_10_names')
)
server <- function(input, output, session){
  top_10_names <- function(){
    babynames::babynames %>% 
      filter(sex == input$sex) %>% 
      filter(year == input$year) %>% 
      top_n(10, prop)
  }
  # MODIFY CODE BELOW: Render a DT output named "table_top_10_names"
  output$table_top_10_names <- DT::renderDT({
    top_10_names()
  })
}
shinyApp(ui = ui, server = server)

```

### 2.5) Add interactive plot output

```{r}

top_trendy_names <- babynames::babynames %>%
                    group_by(name) %>%
                    summarise(Total = sum(n)) %>%
                    dplyr::top_n(10) %>%
                    arrange(name) %>%
                    print()

```


```{r}

library(shiny)
library(ggplot2)

ui <- fluidPage(
  selectInput('name', 'Select Name', top_trendy_names$name),
  # CODE BELOW: Add a plotly output named 'plot_trendy_names'
  plotly::plotlyOutput("plot_trendy_names")
  
)
server <- function(input, output, session){
  # Function to plot trends in a name
  plot_trends <- function(){
     babynames::babynames %>% 
      filter(name == input$name) %>% 
      ggplot(aes(x = year, y = n)) +
      geom_col()
  }
  # CODE BELOW: Render a plotly output named 'plot_trendy_names'
  output$plot_trendy_names <- plotly::renderPlotly({
    plot_trends()
  })

  
  
}
shinyApp(ui = ui, server = server)

```

### 2.6) Sidebar layouts

```{r}

ui <- fluidPage(
  # MODIFY CODE BELOW: Wrap in a sidebarLayout
        sidebarLayout(
          # MODIFY CODE BELOW: Wrap in a sidebarPanel
          sidebarPanel(
            selectInput('name', 'Select Name', top_trendy_names$name)),
            # MODIFY CODE BELOW: Wrap in a mainPanel
            mainPanel(
              plotly::plotlyOutput('plot_trendy_names'),
              DT::DTOutput('table_trendy_names')
      )
    )
)
# DO NOT MODIFY
server <- function(input, output, session){
  # Function to plot trends in a name
  plot_trends <- function(){
     babynames::babynames %>% 
      filter(name == input$name) %>% 
      ggplot(aes(x = year, y = n)) +
      geom_col()
  }
  output$plot_trendy_names <- plotly::renderPlotly({
    plot_trends()
  })
  
  output$table_trendy_names <- DT::renderDT({
    babynames::babynames %>% 
      filter(name == input$name)
  })
}
shinyApp(ui = ui, server = server)

```

### 2.7) Tab layouts

```{r}

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput('name', 'Select Name', top_trendy_names$name)
    ),
    mainPanel(
      # MODIFY CODE BLOCK BELOW: Wrap in a tabsetPanel
      tabsetPanel(
        # MODIFY CODE BELOW: Wrap in a tabPanel providing an appropriate label
        tabPanel("Plot",
          plotly::plotlyOutput('plot_trendy_names')
        ),
        # MODIFY CODE BELOW: Wrap in a tabPanel providing an appropriate label
        tabPanel("Table",
        DT::DTOutput('table_trendy_names')
        )
      )
    )
  )
)
server <- function(input, output, session){
  # Function to plot trends in a name
  plot_trends <- function(){
     babynames::babynames %>% 
      filter(name == input$name) %>% 
      ggplot(aes(x = year, y = n)) +
      geom_col()
  }
  output$plot_trendy_names <- plotly::renderPlotly({
    plot_trends()
  })
  
  output$table_trendy_names <- DT::renderDT({
    babynames::babynames %>% 
      filter(name == input$name)
  })
}
shinyApp(ui = ui, server = server)

```

### 2.8) Themes


```{r}

library(shiny)
library(tidyverse)
library(shinythemes)

ui <- fluidPage(
  # CODE BELOW: Add a titlePanel with an appropriate title
  titlePanel("Teste"),
  
  # REPLACE CODE BELOW: with theme = shinythemes::shinytheme("<your theme>")
  shinythemes::shinytheme('superhero'),
  sidebarLayout(
    sidebarPanel(
      selectInput('name', 'Select Name', top_trendy_names$name)
    ),
    mainPanel(
      tabsetPanel(
        tabPanel('Plot', plotly::plotlyOutput('plot_trendy_names')),
        tabPanel('Table', DT::DTOutput('table_trendy_names'))
      )
    )
  )
)
server <- function(input, output, session){
  # Function to plot trends in a name
  plot_trends <- function(){
     babynames::babynames %>% 
      filter(name == input$name) %>% 
      ggplot(aes(x = year, y = n)) +
      geom_col()
  }
  output$plot_trendy_names <- plotly::renderPlotly({
    plot_trends()
  })
  
  output$table_trendy_names <- DT::renderDT({
    babynames::babynames %>% 
      filter(name == input$name)
  })
}
shinyApp(ui = ui, server = server)

```
### 2.9) App 1: Multilingual Greeting

```{r}

library(shiny)

ui <- fluidPage(
  
  titlePanel('Select Greeting'),
  selectInput('greeting', 'Select greeting', choices = c('Hello', 'Bonjour')),
  textInput('name', 'Enter your name'),
  textOutput('greeting')
  
)

server <- function(input, output, session){
  
  output$greeting <- renderText({
    paste(input$greeting, input$name)
  })
  
}

shinyApp(ui = ui, server = server)
```

### 2.10) App 2: Popular Baby Names

```{r}

get_top_names <- function(year, sex){
  babynames::babynames %>%
    filter(year == year) %>%
    filter(sex == sex) %>%
    dplyr::top_n(10) %>%
    mutate(name = forcats::fct_inorder(name))
}

```

```{r}

library(shiny)
library(dplyr)
library(ggplot2)

ui <- fluidPage(

    titlePanel('Most Popular Names'),
    sidebarLayout(
        sidebarPanel(
            selectInput('sex', 'Select sex', choices = c('M', 'F')),
            sliderInput('year', 'Select Year', value = 1880, min = 1880, max = 2017)
        ),
        mainPanel(
            plotOutput('plot')
        )
    )

)

server <- function(input, output, session) {
    output$plot <- renderPlot({
        top_names_by_sex_year <- get_top_names(input$sex, input$year)
        ggplot(top_names_by_sex_year, aes(x = year, y = sex)) +
        geom_col()
    }) 
}
shinyApp(ui = ui, server = server)

```
### 2.11) App 3: Popular Baby Names Redux

```{r}

ui <- fluidPage(
  titlePanel("Most Popular Names"),
  sidebarLayout(
    sidebarPanel(
      selectInput('sex', 'Select Sex', c("M", "F")),
      sliderInput('year', 'Select Year', 1880, 2013, 1900)
    ),
    mainPanel(
      # MODIFY CODE BLOCK BELOW: Wrap in a tabsetPanel
      tabsetPanel(
        # MODIFY CODE BELOW: Wrap in a tabPanel providing an appropriate label
        tabPanel("Plot",
                 plotOutput('plot')
        ),
        # MODIFY CODE BELOW: Wrap in a tabPanel providing an appropriate label
        tabPanel("Table",
                 tableOutput('table')
        )
      )
    )
  )
)
server <- function(input, output, session) {
  output$plot <- renderPlot({
    top_names_by_sex_year <- get_top_names(input$year, input$sex) 
    ggplot(top_names_by_sex_year, aes(x = name, y = prop)) +
      geom_col()
  })
  
  output$table <- renderTable({
    get_top_names(input$year, input$sex) 
    
  })
}
shinyApp(ui = ui, server = server)

```
## Reactive programming

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel('BMI Calculator'),
  sidebarLayout(
    sidebarPanel(
      numericInput('height', 'Enter your height in meters', 1.5, 1, 2),
      numericInput('weight', 'Enter your weight in Kilograms', 60, 45, 120)
    ),
    mainPanel(
      textOutput("bmi"),
      textOutput("bmi_range")
    )
  )
)

server <- function(input, output, session) {
  # CODE BELOW: Add a reactive expression rval_bmi to calculate BMI
   rval_bmi <- reactive({
    input$weight/(input$height^2)
    })
  output$bmi <- renderText({
    # MODIFY CODE BELOW: Replace right-hand-side with reactive expression
    bmi <- rval_bmi()
    paste("Your BMI is", round(bmi, 1))
  })
  output$bmi_range <- renderText({
    # MODIFY CODE BELOW: Replace right-hand-side with reactive expression
    bmi <- rval_bmi()
    bmi_status <- cut(bmi, 
      breaks = c(0, 18.5, 24.9, 29.9, 40),
      labels = c('underweight', 'healthy', 'overweight', 'obese')
    )
    paste("You are", bmi_status)
  })
}


shinyApp(ui = ui, server = server)

```

```{r}

server <- function(input, output, session) {
  rval_bmi <- reactive({
    input$weight/(input$height^2)
  })
  # CODE BELOW: Add a reactive expression rval_bmi_status to 
  # return health status as underweight etc. based on inputs
  rval_bmi_status <- reactive({
    cut(rval_bmi(), 
      breaks = c(0, 18.5, 24.9, 29.9, 40),
      labels = c('underweight', 'healthy', 'overweight', 'obese')
    )
  })
  output$bmi <- renderText({
    bmi <- rval_bmi()
    paste("Your BMI is", round(bmi, 1))
  })
  output$bmi_status <- renderText({
    # MODIFY CODE BELOW: Replace right-hand-side with 
    # reactive expression rval_bmi_status
    bmi_status <- rval_bmi_status()
    paste("You are", bmi_status)
  })
}
ui <- fluidPage(
  titlePanel('BMI Calculator'),
  sidebarLayout(
    sidebarPanel(
      numericInput('height', 'Enter your height in meters', 1.5, 1, 2),
      numericInput('weight', 'Enter your weight in Kilograms', 60, 45, 120)
    ),
    mainPanel(
      textOutput("bmi"),
      textOutput("bmi_status")
    )
  )
)

shinyApp(ui = ui, server = server)

```
```{r}

ui <- fluidPage(
  textInput('name', 'Enter your name')
)

server <- function(input, output, session) {
  # CODE BELOW: Add an observer to display a notification
  # 'You have entered the name xxxx' where xxxx is the name
  observe({
    showNotification(
      paste("You have entered the name", input$name)
    )
  })
}

shinyApp(ui = ui, server = server)

```
```{r}

server <- function(input, output, session) {
  rval_bmi <- reactive({
    input$weight/(input$height^2)
  })
  output$bmi <- renderText({
    bmi <- rval_bmi()
    # MODIFY CODE BELOW: 
    # Use isolate to stop output from updating when name changes.
    paste("Hi", isolate({input$name}), ". Your BMI is", round(bmi, 1))
  })
}
ui <- fluidPage(
  titlePanel('BMI Calculator'),
  sidebarLayout(
    sidebarPanel(
      textInput('name', 'Enter your name'),
      numericInput('height', 'Enter your height (in m)', 1.5, 1, 2, step = 0.1),
      numericInput('weight', 'Enter your weight (in Kg)', 60, 45, 120)
    ),
    mainPanel(
      textOutput("bmi")
    )
  )
)

shinyApp(ui = ui, server = server)

```

```{r}

server <- function(input, output, session) {
  # MODIFY CODE BELOW: Use eventReactive to delay the execution of the
  # calculation until the user clicks on the show_bmi button (Show BMI)
  rval_bmi <- eventReactive(input$show_bmi, {
    input$weight/(input$height^2)
  })
  output$bmi <- renderText({
    bmi <- rval_bmi()
    paste("Hi", input$name, ". Your BMI is", round(bmi, 1))
  })
}
ui <- fluidPage(
  titlePanel('BMI Calculator'),
  sidebarLayout(
    sidebarPanel(
      textInput('name', 'Enter your name'),
      numericInput('height', 'Enter height (in m)', 1.5, 1, 2, step = 0.1),
      numericInput('weight', 'Enter weight (in Kg)', 60, 45, 120),
      actionButton("show_bmi", "Show BMI")
    ),
    mainPanel(
      textOutput("bmi")
    )
  )
)


shinyApp(ui = ui, server = server)

```
```{r}

server <- function(input, output, session) {
  # MODIFY CODE BELOW: Wrap in observeEvent() so the help text 
  # is displayed when a user clicks on the Help button.
  observeEvent(input$show_help, {
     # Display a modal dialog with bmi_help_text
     # MODIFY CODE BELOW: Uncomment code
     showModal(modalDialog(bmi_help_text))
  })
  rv_bmi <- eventReactive(input$show_bmi, {
    input$weight/(input$height^2)
  })
  output$bmi <- renderText({
    bmi <- rv_bmi()
    paste("Hi", input$name, ". Your BMI is", round(bmi, 1))
  })
}

ui <- fluidPage(
  titlePanel('BMI Calculator'),
  sidebarLayout(
    sidebarPanel(
      textInput('name', 'Enter your name'),
      numericInput('height', 'Enter your height in meters', 1.5, 1, 2),
      numericInput('weight', 'Enter your weight in Kilograms', 60, 45, 120),
      actionButton("show_bmi", "Show BMI"),
      # CODE BELOW: Add an action button named "show_help"
      actionButton("show_help", "Help")
      
    ),
    mainPanel(
      textOutput("bmi")
    )
  )
)

shinyApp(ui = ui, server = server)

```

```{r}

server <- function(input, output, session) {
  # MODIFY CODE BELOW: Delay the height calculation until
  # the show button is pressed
  rval_height_cm <- eventReactive(input$show_height_cm, {
    input$height * 2.54
  })
  
  output$height_cm <- renderText({
    height_cm <- rval_height_cm()
    	paste("Your height in centimeters is", height_cm, "cm")
    })
}

ui <- fluidPage(
  titlePanel("Inches to Centimeters Conversion"),
  sidebarLayout(
    sidebarPanel(
      numericInput("height", "Height (in)", 60),
      actionButton("show_height_cm", "Show height in cm")
    ),
    mainPanel(
      textOutput("height_cm")
    )
  )
)

shinyApp(ui = ui, server = server)

```

```{r}

server <- function(input, output, session) {
  # MODIFY CODE BELOW: Delay the height calculation until
  # the show button is pressed
  rval_height_cm <- eventReactive(input$show_height_cm, {
    input$height * 2.54
  })
  
  output$height_cm <- renderText({
    height_cm <- rval_height_cm()
    	paste("Your height in centimeters is", height_cm, "cm")
    })
}

ui <- fluidPage(
  titlePanel("Inches to Centimeters Conversion"),
  sidebarLayout(
    sidebarPanel(
      numericInput("height", "Height (in)", 60),
      actionButton("show_height_cm", "Show height in cm")
    ),
    mainPanel(
      textOutput("height_cm")
    )
  )
)

shinyApp(ui = ui, server = server)

```

